SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source Term Evaluation

### 2018-02-20 paper.tex
edits

\subsection{Performance}
\label{S:results}
The performance studies in this work were run on the platforms listed in~\cref{t:cpus,t:gpus}.
Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc 4.8.5} on the \sse/\slash\gpuold/ machines.
The optimization level ``\texttt{-O3}'' was used in all cases, and no ``fast-math'' OpenCL optimizations were enabled.
Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.
Finally, unless otherwise stated the \conp/ assumption was used, a vector width of \num{8}\slash\num{128} was used for all CPU\slash GPU cases respectively, ``C''\slash ``F''-ordered data was used on the CPU\slash GPU, and the runtimes reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
The effects of choice of vector-width on the CPU\slash GPU, as well as the differences between \conp/ and \conv/ evaluations will be explored in~\cref{S:source_results}, and briefly discussed in~\cref{S:jacobian_results}.

@@

\Cref{F:cpu_source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
Source term evaluations, critial in their own right for direct numerical simulations~\cite{Spafford:2010aa} among other applications, also give us a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation peformance.

In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
As expected, the \avx/ machine is moderately faster than the \sse/ CPU; additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.

In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
In contrast,~\cref{F:pocl_source} shows the runtime of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work, however briefly speaking vectorization might be achieved on POCL via a combination of relaxed math (``-cl-fast-relaxed-math''), use of explicit OpenCL vector types and newer versions of LLVM (specifically, potential improvments to the vectorization module).
This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.

\begin{figure}[htbp!]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied.}
      \label{F:intel_source_noorm}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{intel_source.pdf}
      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and shallow-vectorized Intel OpenCL codes; the speedup is presented on per-machine basis, e.g., the \avx2/ shallow-vectorized code speedup is relative to OpenMP on the same CPU.}
      \label{F:intel_source}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization.}
      \label{F:pocl_source}
  \end{subfigure}
 \caption{Mean runtimes per condition and speedups achieved by the various CPU OpenCL runtimes compared to OpenMP parallelization for each chemical model studied. The names in the legends correspond to the identifiers listed in~\cref{t:cpus}.}
 \label{F:cpu_source}
\end{figure}

\Cref{F:gpu_source} shows the performance of the source term evaluation on the GPUs listed in~\cref{t:gpus}.
In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease for all chemical models until \textasciitilde\num{e4} conditions, at which point the GPU becomes saturated and performance levels off.
\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling}; the speedup ranges from \textasciitilde\SIrange{1.75}{2.5}{$\times$} depending on the chemical model in question.
Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width size (a.k.a., the block size in NVIDIA terminology) has minimal effect on performance; it is possible that use of larger vector sizes could have an effect on performance, but this was not considered in this work.\todo{rerun NVIDIA on srv2 -- see if that high variability in isopentanol persists}

@@

\Cref{f:source_permutate} shows the effect of changing data-ordering patterns as well as the \conp/ or \conv/ formulation on the performance of source term evaluation in \texttt{pyJac}.
In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on runtime for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
Generally speaking, the difference between the \conp/ and \conv/ formulations has little effect on performance regardless of CPU\slash GPU choice, or even for Jacobian evaluation which will be explored later.

In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-order OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
Additionally, we note that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
In general, the data-ordering has a much more profound affect on the vectorized-OpenCL codes; this is due to the nature of SIMD-operations, which are much more efficient when operating on contiguous (in memory) data rather than incurring an expensive memory ``gather'' operation.
The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \textasciitilde\SIrange{1.5}{2.5}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered shallow-vectorized OpenCL implementation.
However, the ``C''-ordered shallow-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.

\Cref{F:source_gpu_cvsf} shows the effect of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
The speedup reported is calculated per-GPU, such that the effect of data ordering can be seen on each individual GPU.
In general, the ``C''-ordered shallow-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case). \todo{again, pending rerun on \gpuold/}
The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality; coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
      \caption{The mean runtime per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP for \avx/.}
      \label{F:source_conpvsconv}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_cvsf.pdf}
      \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx/ CPU.  The speedup presented is calculated per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
      \label{F:source_cvsf}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_vector_width.pdf}
      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
      \label{F:source_vector_width}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector width of 128.  The speedup presented is calculated per-GPU to better assess the effect of the data-ordering on each GPU.}
      \label{F:source_gpu_cvsf}
  \end{subfigure}
  \caption{The effect of the \conp/ and \conv/ formulations, as well as ``C'' and ``F'' data-ordering on source term evaluation performance in \texttt{pyJac}.
	   It is noted the shallow-vectorized ``C'' ordered OpenCL case corresponds to the vectorized data ordering described in~\cref{S:data}}
  \label{f:source_permutate}
\end{figure}

@@

\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (and in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde$\SI{2}{\sec}$ for OpenCL, and \textasciitilde$\SI{9}{\sec}$ for OpenMP.
By the time 4--8 CPU cores are utilized, the runtime of shallow-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
In contrast, the OpenMP code requires \textasciitilde\num{4} cores to run as fast as the shallow-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
Indeed, we will see that the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.

The SIMD efficiency is defined as:
\begin{equation}
 \label{e:simd_efficiency}
 \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{shallow}}}
\end{equation}
where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}
This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency, \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \textasciitilde\SIrange{10}{25}{\percent} improvement over the nominal vector width.

>>>>

\subsection{Performance}
\label{S:results}
The performance studies in this work were run on the platforms listed in~\cref{t:cpus,t:gpus}.
Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx/\slash\gpunew/ platforms, and \texttt{gcc v4.8.5} on the \sse/\slash\gpuold/ machines.
The optimization level ``\texttt{-O3 -mtune=native}''\todo[inline]{again, rerun OpenMP w/ this} was used, and no ``fast-math'' OpenCL optimizations were enabled.
Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.
Finally, unless stated otherwise, the \conp/ assumption was used, along with a vector width of \num{8}\slash\num{128} for all CPU\slash GPU cases respectively; ``C''\slash ``F''-ordered data was used on the CPU\slash GPU, and the runtimes reported are for the number of conditions specified in~\cref{T:error} and include data transfer overhead to\slash from internal buffers used in \texttt{pyJac}.
The effects of choice of vector-width on the CPU\slash GPU, as well as the differences between \conp/ and \conv/ evaluations will be explored in~\cref{S:source_results}, and briefly discussed in~\cref{S:jacobian_results}.

@@

\Cref{F:cpu_source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
Source term evaluations---critial in their own right for direct numerical simulations of reacting flow~\cite{Spafford:2010aa}, among other applications---also give us a convenient platform to detail the effects of various code configuration options before investigating the more involved Jacobian evaluation peformance.

In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx/ and~\sse/ CPUs, using both Intel OpenCL and OpenMP.
This normalization of the runtime by the number of initial conditions tested is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}).
For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the unvectorized (strictly parallely) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
As expected, the \avx/ machine is moderately faster than the \sse/ CPU for these strictly parallely cases; additionally, the shallow-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.

In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is calculated per-machine, e.g., the \avx/ shallow-vectorized code speedup is relative to OpenMP on the same CPU such that the speedup of the shallow-vectorized OpenCL code can be readily assessed.
On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
Additionally, the shallow-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse/ and \avx/ machines respectively.
In contrast,~\cref{F:pocl_source} shows the runtime of deep, shallow and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work; however, briefly speaking vectorization might be achieved on POCL via a combination of relaxed math (``\texttt{-cl-fast-relaxed-math}''), use of explicit OpenCL vector types and newer versions of LLVM\slash POCL (specifically, potential improvements to the vectorization module).
This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.

\begin{figure}[htbp!]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied.}
      \label{F:intel_source_noorm}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{intel_source.pdf}
      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and shallow-vectorized Intel OpenCL codes; the speedup is presented on per-machine basis.}
      \label{F:intel_source}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization.}
      \label{F:pocl_source}
  \end{subfigure}
 \caption{Mean runtimes per condition and speedups achieved by the various CPU OpenCL runtimes compared to OpenMP parallelization for each chemical model studied. The names in the legends correspond to the identifiers listed in~\cref{t:cpus}.}
 \label{F:cpu_source}
\end{figure}

\Cref{F:gpu_source} shows the performance of the source term evaluation on the GPUs listed in~\cref{t:gpus}.
In~\cref{F:gpu_source_scaling}, the number of initial conditions over which the source terms are evaluated on the \gpunew/ GPU is varied in order to determine the effect on the mean runtime per condition; the mean runtime continues to decrease until \textasciitilde\num{e4} conditions for all chemical models (except the \ce{H2}\slash\ce{CO} model, which is closer to \textasciitilde\num{3e4} conditions), at which point the GPU becomes saturated and performance levels off.
\Cref{F:gpu_source_speedup} shows the speedup in source term evaluations the \gpunew/ GPU over the \gpuold/ GPU for the maximum number of conditions in~\cref{F:gpu_source_scaling}; the speedup ranges from \textasciitilde\SIrange{1.75}{2.5}{$\times$} depending on the chemical model in question.
Additionally,~\cref{F:gpu_source_speedup} shows that varying the vector width size (a.k.a., the block size in NVIDIA terminology) has minimal effect on performance; it is possible that use of larger vector sizes could have an effect on performance, but this was not considered in this work.\todo{rerun NVIDIA on srv2 -- see if that high variability in isopentanol persists}

@@

\Cref{f:source_permutate} shows the effect of changing data-ordering patterns, the \conp/ or \conv/ formulation, and the CPU vector-width size on the performance of source term evaluation in \texttt{pyJac}.
In~\cref{F:source_conpvsconv}, we see that choice of the constant-pressure or constant-volume formulation has little to no effect on runtime for OpenMP as well as the shallow-vectorized\slash unvectorized Intel OpenCL codes on the \avx/ machine.
Generally speaking, the difference between the \conp/ and \conv/ formulations has little effect on performance regardless of CPU\slash GPU choice, or even for Jacobian evaluation which will be explored later.

In contrast,~\Cref{F:source_cvsf} shows large speedups of ``C''-ordered data over ``F''-ordered data on the \avx/ machine; the speedup presented is calculated per language, i.e., the roughly \SIrange{1.5}{2}{$\times$} speedup of the ``C''-ordered OpenMP implementation is relative to the ``F''-ordered OpenMP baseline while the OpenCL speedups are given with respect to the ``F''-ordered unvectorized OpenCL code.
Additionally, we note that the ``C''-ordered shallow-vectorization in~\cref{F:source_cvsf} utilizes the vectorized-data ordering described in~\cref{S:data}.
In general, the data-ordering has a much more profound affect on the vectorized-OpenCL codes; this is due to the nature of SIMD-operations, which are much more efficient when operating on contiguous (in memory) data rather than incurring an expensive memory ``gather'' operation.
The ``C''-ordered unvectorized OpenCL code (which uses the default unvectorized row-major data ordering) is \textasciitilde\SIrange{1.5}{2.5}{$\times$} faster than the unvectorized ``F''-ordered code, and is in some cases almost as fast as the ``F''-ordered shallow-vectorized OpenCL implementation.
However, the ``C''-ordered shallow-vectorization achieves speedups of \textasciitilde\SIrange{4}{7}{$\times$} demonstrating the value of the vectorized data ordering for CPU execution.

\Cref{F:source_gpu_cvsf} shows the mean runtime per-condition of the ``C'' versus ``F''-ordering on source term evaluation performance on both GPUs.
In general, the ``C''-ordered shallow-vectorization is \textasciitilde\SIrange{1.05}{1.2}{$\times$} faster on the \gpunew/ GPU while the ``C'' and ``F''-ordered data are nearly equivalent for the \gpuold/ GPU, except for the isopentanol model where the ``F''-ordered data is \textasciitilde\SI{1.4}{$\times$} faster (though, we note that there was high variablity in this case). \todo{again, pending rerun on \gpuold/}
The ``C''-ordered speedup on the \gpunew/ GPU is counter to what one might expect, as typically speaking coalesced memory access in a shallow-vectorization is easier to achieve if ``F''-ordering is used (see~\cref{S:data}).
However, the vectorized-data ordering here ensures that not only are memory accesses aligned to the vector width size (and thus, likely coalesced) but the improved data-locality, coupled with the larger cache sizes of the \gpunew/ GPU result in this modest speedup for ``C''-data ordering.

\Cref{F:source_vector_width} shows the effect of changing vector-width size on source rate evaluation performance the \avx/ CPU.
The vector-width of size \num{8} was the fastest tested, while the larger vector-width of \num{16} is slightly slower due to increased register pressure~\cite{intel_vectypes}.
It is unclear why the vector-width of \num{4} results in no speedup at all (it is in fact, the slowest case); Intel's vectorization guide~\cite{intel_vecknobs} mentions that a heuristic determines the optimal vector-width (in this case, it appears from compiler output to be \num{8}), so it is possible using vector-width smaller than the heuristic breaks the implicit vectorizer.
It is noted that this issue does not occur for a vector width of \num{4} on the \sse/ CPU.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_conpvsconv.pdf}
      \caption{The mean runtime per condition for each chemical model using both the \conp/ and \conv/ formulations on Intel OpenCL and OpenMP for \avx/.}
      \label{F:source_conpvsconv}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_cvsf.pdf}
      \caption{The speedup achieved by ``C'' ordering over ``F'' ordering for Intel OpenCL and OpenMP on the \avx/ CPU.  The speedup presented is calculated per-language (OpenMP and OpenCL) to better assess the effect of the data-ordering.}
      \label{F:source_cvsf}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_gpu_cvsf.pdf}
      \caption{The affect on performance of ``C'' vs ``F''-ordering for the shallow-vectorized NVIDIA OpenCL code on both GPUs with a vector width of 128.  The speedup presented is calculated per-GPU to better assess the effect of the data-ordering on each GPU.}
      \label{F:source_gpu_cvsf}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{source_vector_width.pdf}
      \caption{The effect of vector-width on ``C''-ordered shallow-vectorized Intel OpenCL source term evaluations in the \avx/ CPU.}
      \label{F:source_vector_width}
  \end{subfigure}
  \caption{The effect of the \conp/ and \conv/ formulations, ``C'' and ``F'' data-ordering, and CPU vector-width on source term evaluation performance in \texttt{pyJac}.
	   It is noted the shallow-vectorized ``C'' ordered OpenCL cases corresponds to the vectorized data ordering described in~\cref{S:data}}
  \label{f:source_permutate}
\end{figure}

@@

\Cref{F:source_parallel_scaling} shows the strong parallel scaling efficiency of source rate evaluation in \texttt{pyJac} on the \avx/ machine for both the shallow-vectorized Intel OpenCL and OpenMP codes.
In general, the \ce{H2}\slash\ce{CO} mechanism has the worst scaling efficiency for Intel OpenCL; although the mechanism scalings vary somewhat for OpenMP, there is no clear trend.
It is likely the poor scaling of the \ce{H2}\slash\ce{CO} for Intel OpenCL results from both its relatively small size, and few falloff\slash chemically activated reaction (and in particular, the additional expensive logarithm and exponential evaluations that accompany them); i.e., there is simply not enough work to fully utilize the SIMD intructions.
Additionally, OpenMP tends to have higher scalings then the shallow-vectorized OpenCL code, e.g., \textasciitilde\num{0.8} and \textasciitilde\num{0.7} for \num{4} and \num{8} CPU cores respectively, compared to just \textasciitilde\num{0.7} and \textasciitilde\num{0.5} for OpenCL; although not pictured (to keep the figure readable), the unvectorized Intel OpenCL code has only slightly worse scaling than the OpenMP code, hence the SIMD-vectorization itself the cause of this poorer scaling.
This is again likely due to relatively small amounts of work associated with source rate evaluation; the longest runtime (the isopentanol model on a single core) takes only \textasciitilde\SI{2}{\second} for OpenCL, and \textasciitilde\SI{9}{\second} for OpenMP.
By the time 4--8 CPU cores are utilized, the runtime of shallow-vectorized OpenCL code can be measured in hundreds of milliseconds, and parallel overhead begins to dominate, decreasing efficiency.
In contrast, the OpenMP code requires \textasciitilde\num{4} cores to match the performance of the shallow-vectorized OpenCL code in this case, hence parallel overhead concerns become increasingly important for 8--16 cores.
Indeed, as will be investigated in~\cref{s:jacobian_results} the strong scalings for sparse Jacobian evaluation---the most computationally intensive task in this work for which the efficiency will investigated---are far more similar between Intel OpenCL and OpenMP.

The SIMD efficiency is defined as:
\begin{equation}
 \label{e:simd_efficiency}
 \varepsilon_{\text{SIMD}} = \frac{\bar{t}_{\text{unvec}}}{W \bar{t}_{\text{shallow}}}
\end{equation}
where $\bar{t}_{\text{unvec}}$ is the mean runtime per condition of the unvectorized OpenCL code, $\bar{t}_{\text{shallow}}$ the same for the shallow-vectorized OpenCL code, and $W$, the vector width reported in number of double operations in~\cref{t:cpus}.
This measure compares the actual speedup due to shallow-vectorization compared to the nominal vector-width of the machine in question.
\Cref{F:source_simd_scaling} shows the SIMD efficiency of source term evaluation in \texttt{pyJac} on a single core of the \avx/ machine; the larger models have higher SIMD-efficiency of \textasciitilde\num{0.75}, and the smaller models have lower SIMD-efficiency, \textasciitilde\num{0.625}.
This again demonstrates that the SIMD-vectorization becomes efficient as the amount of work to perform increases (with increasing model size).
Interestingly,~\cref{F:source_sse_simd_scaling} shows the SIMD efficiency on the \sse/ machine as greater than one; this is likely caused by a combination of use of an OpenCL vector-width greater than the native CPU vector-width (i.e., \num{8} v.s., \num{2}), and the improved data-locality for the vectorized-data ordering, as discussed in~\cref{S:data}, and results a modest \textasciitilde\SIrange{10}{25}{\percent} improvement over the nominal vector width.
